---
title: "RangeX project"
author: "Timothée Darbois"
date: "2024-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ade4)
library(ggplot2)
library(tidyverse)
library(googlesheets4)
library(dplyr)
library(rstatix)

library(sjPlot)
library(contrast)
```

```{r}
df<-read_sheet('https://docs.google.com/spreadsheets/d/1EC8b0MkNI7F2d2fFpy_64ZuBGXhM180vjE66Cq11lJQ/edit#gid=0')
df<-as.data.frame(df)
df1<- data.frame(df,student='Timothée')

```

data cleaning
```{r}
df1<-df1 %>%
  filter(!is.na(flowers))
df1<-df1%>%
  filter(!is.na(weight.inf))
df1<-data.frame(df1,condi=paste(df1$site,df1$plotID,sep=" "))

```

see how much data I have per species


```{r}
species=c("CC","CN","SP","TP","PS","PL","LM","SD","HM","LV")
select_species=c("CN","SP","TP","PS","PL","LM","LV") # with enough counts

#df_TP<-df1[df1$species=="TP",]
counts<-c()
for (i in species){
  print(i)
  counts<-c(counts,nrow(df1[df1$species==i,]))

}
count_data<-data.frame(species,counts)



ggplot(data=count_data,aes(x=species,y=counts)) +geom_bar(stat="identity") + ggtitle("timothee data")


```
simple boxplot: effect of treatments on seeds


## import of ingrid's data
```{r}
dff<-read_sheet('https://docs.google.com/spreadsheets/d/1PyHHmGMpWnRUsCRoC8_AoNDIC8ap8zlRJ8HwA1HqLEQ/edit#gid=447429113',na = c("","NA","NULL",NULL))
dff<-as.data.frame(dff)
df2<- dff
df2<-data.frame(df2,student='Ingrid')
```
```{r}
df2<-df2 %>%
  filter(!is.na(flowers))
df2<-df2%>%
  filter(!is.na(weight.inf))

species=c("CC","CN","SP","TP","PS","PL","LM","SD","HM","LV")
#df_TP<-df1[df1$species=="TP",]
counts<-c()
for (i in species){
  print(i)
  counts<-c(counts,nrow(df2[df2$species==i,]))

}
count_data<-data.frame(species,counts)



ggplot(data=count_data,aes(x=species,y=counts)) +geom_bar(stat="identity") +ggtitle('data_ingrid')

```
```{r}
df3<-df1%>% select(1,2,3,4,5,6,7,8,9,10,11,12,13,"student")
df4<-df2%>% select(1,2,3,4,5,6,7,8,9,10,11,12,13,"student")
df5<- rbind(df3,df4)
df5<-data.frame(df5,condi=paste(df5$site,df5$plotID,sep=" "))
df5<-df5 %>%
  filter(!is.na(flowers))
df5<-df5%>%
  filter(!is.na(weight.inf))


```


for each species, seed number
```{r}

plot_snb<- ggplot(df5, aes(x=condi, y=seeds_number.inf))+geom_boxplot(aes(fill=student)) + geom_dotplot(binaxis = 'y',
        dotsize = 1,
        stackdir = 'center') +
    theme_classic() + facet_wrap(vars(species),nrow=2,scales='free')+ggtitle("seed number per infructescence")
  

plot_snb
```
for each species, seed weight
```{r}

plot_sw<- ggplot(df5, aes(x=condi, y=weight.inf))+geom_boxplot(aes(fill=student)) + geom_dotplot(binaxis = 'y',
        dotsize = 1,
        stackdir = 'center') +
    theme_classic() + facet_wrap(vars(species),nrow=2,scales='free')+ggtitle("seed weight per infructescence")
  

plot_sw
```

for each species, weight per seeds
```{r}
plot_wps<- ggplot(df5, aes(x=condi, y=weight.seed))+geom_boxplot(aes(fill=student)) + geom_dotplot(binaxis = 'y',
        dotsize = 1,
        stackdir = 'center') +
    theme_classic() + facet_wrap(vars(species),nrow=2,scales='free')+ggtitle("seed weight per seed")
  

plot_wps
```

# analyse stat
## effet expérimentateur
```{r}
df5_TP<-df5[df5$species=="TP",]
df5_TP<-df5_TP %>%
  filter(!is.na(weight.inf))


```

## test of normality
```{r}
plot_sw<- ggplot(df5, aes(x= weight.inf))+geom_histogram(aes(fill=student))+
    theme_classic() + facet_wrap(vars(species),nrow=2,scales='free')+ggtitle("seed weight per infructescence")
  
```

## test of normality
```{r}
hist_wpi<- ggplot(df5, aes(x=weight.inf))+
  geom_histogram(aes(fill=student))+
    theme_classic() + 
  facet_wrap(vars(species),nrow=2,scales='free')+
  ggtitle("seed weight per infructescence")
hist_wpi
```


```{r}
shapiro.test(df5$weight.seed)
shapiro.test(log(df5$weight.seed+1))

shapiro.test(df5$seeds_number.inf)

shapiro.test(log(df5[df5$species=="TP",]$weight.inf+1))
shapiro.test(df5[df5$species=="TP",]$weight.inf)


```
at least, some are not normal (eg : there are bimodal distribution with viable and not viable seeds)
#test  effect of the experimentator
```{r}
U_sn<-c()
U_wpi<-c()
U_wps<-c()

for (i in 1:7){
  data_U<-df5[df5$species==select_species[i],]
  
  U_sn<-rbind(U_sn,wilcox.test(seeds_number.inf~student,data_U))

  U_wpi<-rbind(U_wpi,wilcox.test(weight.inf~student,data_U))

  U_wps<-rbind(U_wps,wilcox.test(weight.seed~student,data_U))
}

U_sn<-cbind(U_sn, select_species)
U_wpi<-cbind(U_wpi, select_species)
U_wps<-cbind(U_wps, select_species)

print("seeds number")
U_sn
print("weight.inf")
U_wpi
print("weight per seed")
U_wps
```
## pool all the species together, as part of a community.

```{r}
plot_t_sn<- ggplot(df5, aes(x=condi, y=seeds_number.inf ))+geom_boxplot(aes(fill=student)) + geom_dotplot(binaxis = 'y',
        dotsize = 0.01,
        stackdir = 'center') +theme_classic() 
    
plot_t_sn
anova_test(df5,seeds_number.inf~condi)
anova_test(df5,seeds_number.inf~student)

```

```{r}
plot_t_sw<- ggplot(df5, aes(x=condi, y=log(weight.inf) ))+geom_boxplot(aes(fill=student)) + geom_dotplot(binaxis = 'y',
        dotsize = 0.01,
        stackdir = 'center') +theme_classic() 
    
plot_t_sw
anova_test(df5,weight.inf~condi)
anova_test(df5,weight.inf~student)

```

```{r}
plot_t_wpi<- ggplot(df5, aes(x=condi, y=log(weight.seed) ))+geom_boxplot(aes(fill=student)) + geom_dotplot(binaxis = 'y',
        dotsize = 0.01,
        stackdir = 'center') +theme_classic() 
    
plot_t_wpi
anova_test(df5,weight.seed~condi)
anova_test(df5,weight.inf~student)


```



(what we could have done for anova, looking for interactions with condi.)
```{r}
var=c(8,9,10)

anov_sn<-c()
anov_wpi<-c()
anov_wps<-c()

for (i in 1:7){
  test_anov<-df5[df5$species==select_species[i],]
  
  anov_sn<-rbind(anov_sn,anova_test(test_anov,seeds_number.inf~student*condi))

  anov_wpi<-rbind(anov_wpi,anova_test(test_anov,weight.inf~student*condi))

  anov_wps<-rbind(anov_wps,anova_test(test_anov,weight.seed~student*condi))
}

species_anov<-c()
for (i in 1:7){species_anov<-c(species_anov,rep(select_species[i],3))}
anov_sn<-data.frame(anov_sn,species=species_anov)
anov_wpi<-data.frame(anov_wpi,species=species_anov)
anov_wps<-data.frame(anov_wps,species=species_anov)
print(anov_sn)
print(anov_wpi)
print(anov_wps)

```

## glm


```{r}
glm_allsp_sn <- glm(seeds_number.inf~condi, data=df5,family= "poisson")
glm_allsp_sn
glm_allsp_sn2 <- glm(seeds_number.inf~site, data=df5,family= "poisson")


sjPlot::tab_model(glm_allsp_sn,show.intercept = T)
plot_model(glm_allsp_sn, show.values = TRUE, width = 0.1, title = "effect of treatments on seeds number per infructescence") 
```
let's focus on PL
```{r}
df5_PL<- df5[df5$species== "PL",]
glm_PL_sn <- glm(seeds_number.inf~condi, data=df5_PL,family= "poisson")
summary(glm_PL_sn)
glm_PL_sn2 <- glm(seeds_number.inf~site, data=df5_PL,family= "poisson")
summary(glm_PL_sn2 )

sjPlot::tab_model(glm_PL_sn,show.intercept = T)
plot_model(glm_PL_sn, show.values = TRUE, width = 0.1, title = "effect of treatments on seeds number per infructescence") 
```



glm1 <- glm(Nombre_espece~Topographie, data=tab, family = "poisson")

continue with glmm


#prendre le log.
effet_etude<-lm(weight.inf~student,df5)
shapiro.test(effet_etude$residuals)

#Community Weighted Mean CWM
library(FD)
shapiro.test(effet.etud$residuals)

df5_TP <- df5_TP %>% 
  mutate(
    flowers = replace(flowers, "NULL", NA)
  )
df %>% 
  drop_na(var2)

res.aov <- jobsatisfaction %>% anova_test(score ~ gender * education_level)
res.aov
qqq plot :

qqnorm(scale(body_mass))
abline(a=0, b=1, col = "red")



ggplot(df, aes(x_var, y_var)) +
  geom_point() +
  facet_wrap(vars(category_var))

ggplot(data_air_nona, aes(x = Month, y = Ozone))

box_plot +
    geom_boxplot() +
    geom_dotplot(binaxis = 'y',
        dotsize = 1,
        stackdir = 'center') +
    theme_classic()

ggplot(data=df2, aes(x=dose, y=len, fill=supp)) +
  geom_bar(stat="identity")


facet grid

mixed model : effet de plusieurs variables en interaction sur une variable réponse.: variable fixe et random?
n